The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2018-10-22 Paper.Rmd
adding discussion(primarily) and fixing figures

<!--what we did?-->
In the earlier work, most of the studies in ecology focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of asymmetric tail-dependence and non-Normal copulae in this field. We

<!--why is it important to study?-->

<!--future work-->

>>>>

<!--what we did and why is it important to study?-->

In the earlier work, most of the studies in ecology focused on measuring correlations between variables and thus ignoring the possibility to recognize importance of stronger association between extreme values of them. In the present study, we introduced the concept of asymmetric tail-dependence and non-Normal copulae in this field. We showed by examples of several real data set (Table \ref{tab_bivar_summary} and Table \ref{tab_multivar_summary}) that non-Normal copulae with specific tail dependence is common in nature and could tell us about the nature of interaction between different ecological variables. Our method in this regard of two ways. First, we did a model selection approach based on maximum likelihood which fit the real copula data with an optimal set of 16 copulae family. These copulae were chosen carefully so that they covered the whole range from exclusively lower (Clayton, Survival Gumbel, Survival Joe, Survival BB6, Survival BB8) or upper tail dependence (Survival Clayton, Gumbel, Joe, BB6, BB8) to both tail dependence (BB1, Survival BB1, BB7, Survival BB7) in addition to two symmetric copulae (Normal and Frank). Copula having highest AIC-weight came as best-fit from this approach. Though we did goodness of fit test based on two statistics (Cramer-von Mises and Kolmogorov-Smirnov), still we wanted to make sure the reliability of our model selection results. So, on a second note, we measured the tail dependence at both lower and upper end of a copula by three different non-parametric statistics ($\cor$, $\Ps$ and $\Dsq$) which we innovated and we found they were consistent with the inferences we drew from model selction results. Our approach of "copula" could help to interprete the synchronous behavior of a species from several metapopulations. To illustrate, if pairwise copulae from a multivariate spatial data of a species (e.g. abundance data of green-spruce aphid) showed major "lower-tail dependence" on most of the locations then it could be considered as a signature of "spatial synchrony" that metapopulations are simultaneously rare at those locations. So, if some aphid species showed synchronously "upper-tail dependence" then it would be harmful for corp-production or in case of synchronous abundance of planktons could lead to algal-bloom in the ocean. Copula could also gave us information about the causal relationship between ecological variables. For example, if maximum railfall is beneficial for some species then it would show an upper-tail dependent copula where high values of species abundance and precipitation would be strongerly correlated then lower values.

We explained the possible causes which induced asymmetric tail-dependence in copula. Borrowing the idea from "Moran's theorem", we demonstrated with simulations that two habitat patches when affected by asymmetric bivariate-noise copula showed similar tail-dependence asymmetry in the population copula from those patches (see Fig. \ref{fig_Cause4copula_params_GOF_500} and Fig. \ref{fig_Cause4copula_scatter_C_kend}). We also showed the effect of global environmental fluctuation (behaving asymmetrically beyond a certain threshold) on the meta-population copula which was shaped with similar tail dependence as of the initial global noise copula (Fig. \ref{fig_ms_asym_b_0.1}). We observed significant changes in skewness of spatially averaged timeseries and also in the curvature of spatial Taylor's law for copula having asymmetric tail-dependence than it would have been expected for Normal copula. For example, Fig. \ref{fig_skewness_spearman}(A) showed green spruce aphid's abundance timeseries had right skewed distribution but its skewness was signifiantly lower than it was expected if it would came from its Spearman correlation preserving Normal surrogates. This result indicates less extreme outbreaks of green spruce aphid around UK than it would have been with Normal copula structure. Similarly, methane fluxes showing lower skewness than Normal surrogates mean that the level of green house gas is under-control situation and monitoring this skewness over several decades may be helpful to understand the "green-house effect" on environment.

<!--future work-->
We think this copula approach have scope to investigate into the extinction risks of metapopulation due to climate change. To illustrate, lower-tail dependent environmental fluctuations may suppress the metapopulations synchronously by making some of the habitats go extinct. This could help in managing policy for conservation of endangered populations.
